### 2019-03-25 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
add salmon01, cohortStrength to models

---
title: "Body size patterns"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Plots for the manuscript

```{r Load data, message=FALSE}
library(itsadug)
library(arm)
library(relaimpo)
library(linkedModels)
library(arrayhelpers)
library(mgcv)
library(tidyverse)
library(gridExtra)

#RECALCULATE gr for dataset. Filtering by maxSampleInterval in the creat ddddG step leaves in trailing lengths from the long samp intervals
# d2 <- d2 %>%
#   group_by(tag) %>%
#   # arrange(tag,sampleNumber) %>%
#   mutate( lagObservedWeight = lead(observedWeight),
#           lagObservedLength = lead(observedLength),
#           grWeight = exp(lagObservedWeight - observedWeight)/as.numeric((lagDetectionDate - detectionDate)),
#           grLength = (lagObservedLength - observedLength)/as.numeric((lagDetectionDate - detectionDate))
#   )  %>%
#   ungroup()

load("D:/projects/linkedModels/data/cd_west_1997.RData")
cd <- cd %>% mutate(observedWeight = ifelse(observedWeight == -9999.9, NA, observedWeight),
                    lagObservedWeight = ifelse(lagObservedWeight == -9999.9, NA, lagObservedWeight))
```

```{r gams for body size - data prep}
speciesIn <- 'bkt'
depVarIn <- 'observedLength'
kIn <- 4 # # of knots for the speces and river smooths

vB <- 
  cd %>% mutate( age = year - cohort,
                 metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1),
                 ydayCumul = yday + 365 * (year - cohort),
                 cf = observedWeight/observedLength^3 * 10000 ) %>%
    filter( ageInSamples %in% 1:9,
            age < 3, 
            year < 2016, area %in% c('inside','trib'), enc == 1
            #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
            )

cohortOffset <- 2006

# 1-way
vB$cohortFactor <- factor(vB$cohort)
vB$cohortOrdered <- factor(vB$cohort, levels = 2000:2013, ordered = T)
#vB$cohortFactorCentered <- vB$cohortFactor - 2006
vB$cohortCentered <- vB$cohort - cohortOffset

vB$speciesFactor <- factor(vB$species, levels = c('bkt','bnt'))
vB$speciesOrdered <- factor(vB$species, levels = c('bkt','bnt'), ordered = T)

vB$riverFactor <- factor(vB$river, levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'))
#vB$riverOrdered <- riverOrdered exists

# 2-way
vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
vB$speciesCohortOrdered <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
vB$cohortRiverOrdered <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)

# 3-way
vB$cohortSpeciesRiverOrdered <- factor(paste(vB$cohortOrdered,vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)

vB$tagIndexFactor <- factor(vB$tagIndex)

#vB$riverOrdered <- factor(vB$river,
#                         levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'),
#                         ordered = T)

vB$salmon01 <- ifelse(vB$cohort <=2004, 1, 0)

```

```{r cohort strength}

cohortStrength <- vB %>%
  distinct(tagIndex, species, metaPop01, cohort) %>% #ignoring 10 fish that were in both o'bear and WB
  group_by(species, metaPop01, cohort) %>%
  summarize(cohortStrength = n()) %>%
  filter(!(cohort <2000 & species !='ats')) %>% # no trib data
  mutate(cohortStrengthScaled = scale(cohortStrength))

vB <- left_join(vB, cohortStrength)

ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species)) +
  geom_point(size = 2) +
  geom_line(size = 1.1, color = 'grey40') +
  labs(x = 'Cohort',
       y = 'Cohort strength') +
  theme_publication() +
  facet_grid(~metaPop01)

```

```{r dataIn}
dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
```

Run to here to start

```{r raw data graphs}
# number of knots from gam 
kValue <- 9 # from mod6 below, just for graphing

# all fish
ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength)) +
  geom_point(size = 0.5, alpha = 0.1) +
 # geom_smooth(span = 0.1, size = 1.5, se = F) +
  geom_smooth(method='gam', alpha = 0.1, formula=y~s(x, k = kValue), se = F) +
  theme_publication()

# by species
ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength, color = species)) +
  geom_point(size = 0.5, alpha = 0.1) +
 # geom_smooth(span = 0.1, size = 1.5, se = F) +
  geom_smooth(method='gam', alpha = 0.1, formula=y~s(x, k = kValue), se = F) +
 # facet_grid(~species) +
  theme_publication()

# by species, by cohort gridded
# ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength, group = species, color = (species))) +
#   geom_point(size = 0.5, alpha = 0.1) +
#  # geom_smooth(span = 0.1, size = 1.5, se = F) +
#   geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
#   facet_wrap( ~ cohort) +
#   theme_publication()

# by species (bkt, bnt), by cohort, by river gridded
ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats')), aes(ydayCumul,observedLength, group = species, color = (species))) +
  geom_point(size = 0.5, alpha = 0.1) +
 # geom_smooth(span = 0.1, size = 1.5, se = F) +
  geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
  facet_grid(riverOrdered ~ cohort) +
  theme_publication()
```

```{r gam models - only need to run for model selection}
####################################
# gam models

################################################################################

# base model

mod <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")
summary(mod)

predGrid <- expand.grid(cohort = unique(mod$model$cohort),
                        river = unique(mod$model$river),
                        ydayCumul = seq(250,1000, by = 30))

pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')

predGrid2 <- bind_cols(predGrid,as.data.frame(pred))

ggplot(predGrid2, aes(ydayCumul, pred)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohort) +
  ggtitle('mod')

##########################################################
# by cohortSpeciesRiver

system.time(
modCR <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
             data = dataIn, method = "REML")
)
summary(modCR)
#plot(modCR, pages = 8, scale = 0)

predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predCR <- predict(modCR, newdata = predGridCR)#, type = 'lpmatrix')

predGrid2CR <- bind_cols(predGridCR, as.data.frame(predCR))

ggplot(predGrid2CR, aes(ydayCumul, predCR)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)+
  ggtitle('modCR')

##########################################################
# by cohortSpeciesRiver with s(cohort)

system.time(
modCR_c <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_c)
#plot(modCR_c, pages = 8, scale = 0)

predGridCR_c <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predCR_c <- predict(modCR_c, newdata = predGridCR_c)#, type = 'lpmatrix')

predGrid2CR_c <- bind_cols(predGridCR_c, as.data.frame(predCR_c))

ggplot(predGrid2CR_c, aes(ydayCumul, predCR_c, group = cohort, color = cohort)) +
  geom_point() + 
  geom_line() +
  #ylim(50,210) +
  facet_wrap(~river)+
  ggtitle('modCR_c')

##########################################################
# by cohortRiver - w ind RE

# need more memory
# system.time(
# modCR_Ind <- bam(observedLength ~ cohort * river +
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
#                s(ydayCumul, tagIndexFactor, bs="fs", m=1, k = kIn),
#              data = dataIn, method = "REML")
# )
# summary(modCR_Ind)
# plot(modCR_Ind, pages = 8, scale = 0)

##########################################################
# by cohortSpeciesRiver, additive in s()

system.time(
modC_R <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered), 
             data = dataIn, method = "REML")
)
summary(modC_R)
#plot(modC_R, pages = 4, scale = 0)

predGridC_R <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predC_R <- predict(modC_R, newdata = predGridC_R)#, type = 'lpmatrix')

predGrid2C_R <- bind_cols(predGridC_R,as.data.frame(predC_R))

ggplot(predGrid2C_R, aes(ydayCumul,predC_R)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)+
  ggtitle('modC_R')

##########################

modC <- bam(observedLength ~ cohort + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")

modR <- bam(observedLength ~ river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered), 
             data = dataIn, method = "REML")

##########################################
# add in salmon01 and cohortStrengthScaled

system.time(
modCR_S_c <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
system.time(
modCR_c_cS <- bam(observedLength ~ cohort * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)

system.time(
modCR_S_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_S_c_cS)

system.time(
modCR_SR_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_SR_c_cS)

system.time(
modCR_SR_c_cSR <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered), 
             data = dataIn, method = "REML")
)
summary(modCR_SR_c_cSR)
#plot(modCR_S_c_cS, scale = 0, pages=8)
#plotGamVarBySpecies(var = 's(cohortStrengthScaled)')

AIC(mod,modC,modR,modC_R,modCR, modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR) %>% rownames_to_column() %>% arrange(AIC)

save(mod,modC,modR,modC_R,modCR,modCR_c, modCR_S_c, modCR_c_cS, modCR_S_c_cS, modCR_SR_c_cS, modCR_SR_c_cSR, file = paste0('./data/out/mods_',speciesIn,'.RData'))

```

```{r best models for each species}

# Brook trout

modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

summary(modBKT)
#plot(modBKT, pages = 8, scale = 0)

predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = seq(-2,2, by = 0.1)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predBKT <- predict(modBKT, newdata = predGridBKT)#, type = 'lpmatrix')

predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')

#################################
# Brown trout

modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

summary(modBNT)
#plot(modBNT, pages = 8, scale = 0)

predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = seq(-2,2, by = 0.1)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predBNT <- predict(modBNT, newdata = predGridBNT)#, type = 'lpmatrix')

predGrid2BNT <- bind_cols(predGridBNT, as.data.frame(predBNT)) %>% mutate(pred = predBNT, species = 'bnt')

##################
# combine species' predictions

predGrid2Both <- bind_rows(predGrid2BKT, predGrid2BNT) %>% select(-predBKT, -predBNT)

predGrid2Both$riverGG <- factor(predGrid2Both$riverOrdered,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

save(modBKT,modBNT,predBKT,predBNT,predGrid2Both, file = paste0('./data/out/bestModsBKTBNT.RData'))

```

```{r graphs of best models}
# load('./data/out/bestModsBKTBNT.RData')
source('./functions/plotGams.R')

# cohorts by river/species
ggplot(predGrid2Both %>% filter(cohortStrengthScaled == 0, salmon01 == 0), aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face="plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(speciesGG ~ riverGG, scales = 'free_y')

# specific splines by species
plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort')

# main spline over ydayCumul
plotBKT = plot.gam(modBKT, select = 1, scale = 0)
plotBNT = plot.gam(modBNT, select = 1, scale = 0)

plotBKTydayCumul <- data.frame(x = plotBKT[[1]]$x, y = plotBKT[[1]]$fit, se = plotBKT[[1]]$se) %>% mutate(species = 'bkt')
plotBNTydayCumul <- data.frame(x = plotBNT[[1]]$x, y = plotBNT[[1]]$fit, se = plotBNT[[1]]$se) %>% mutate(species = 'bnt')
plotBothydayCumul <- bind_rows(plotBKTydayCumul,plotBNTydayCumul) %>% mutate(upperCI = y + se * 1.96, lowerCI = y - se * 1.96)

plotBothydayCumul$speciesGG <- factor(plotBothydayCumul$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(plotBothydayCumul, aes(x, group = speciesGG)) + 
  geom_ribbon(aes( ymin = lowerCI, ymax = upperCI, fill  = speciesGG), alpha = 0.5) +
  geom_line(aes(y = y, linetype = speciesGG), size = 1.5) +
  scale_fill_manual(values = c('grey30', 'grey70')) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       linetype = 'Species') +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face="plain"))

# cohort effects

ggplot(tmp2, aes(x,y)) + geom_line()

appraise(modBKT)

```

```{r early gams}
# some early exploration

# 1. Do we need separate splines on cohort in s()? YES, cohortOrdered is best
# 
mod0_0 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_1 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr', by = (cohort)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_2 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr', by = (cohortFactor)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_3 <- bam(observedLength ~ 
                cohort + 
                s(ydayCumul, bs = 'cr') + 
                s(ydayCumul, bs = 'cr', by = (cohortOrdered)), 
              data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_4 <- bam(observedLength ~ cohort + 
                s(ydayCumul, bs = 'cr') + 
                s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)),
                data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

#mod0_4 is best
AIC(mod0_0,mod0_1,mod0_2,mod0_3,mod0_4) %>% rownames_to_column() %>% arrange(AIC)

# 2. structure of smooths
# with ordered factor, need to include an overall smooth, http://www.sfs.uni-tuebingen.de/~jvanrij/Tutorial/GAMM.html 
mod0_4a <- gam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_4b <- gam(observedLength ~ cohortOrdered + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

# mod0_3b is best, but we want to estimate a trend as in mod0_3a
# add + s(ydayCumul, bs = 'cr') to models
AIC(mod0_4,mod0_4a,mod0_4b) %>% rownames_to_column() %>% arrange(AIC)

# 3. gam vs. bam for this big dataset? bam() is MUCH faster!
system.time( mod0_4a_1 <- gam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
# 316 sec
summary(mod0_3a_1)

system.time( mod0_4a_2 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
# 8 sec
summary(mod0_4a_2)

# 4. bs = 'cr', vs default 'tp'. AIC for 'cr' is ~ 300 lower. Stick with 'cr'
mod0_4a_2a <- bam(observedLength ~ cohort + s(ydayCumul) + s(ydayCumul, by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

AIC(mod0_4a_2,mod0_4a_2a)

# 5. Include Individual as a RE??
# Taking a long time, maybe run overnight
vB$tagIndexFactor <- factor(vB$tagIndex)
system.time( mod0_3a_3 <- bam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)) + s(ydayCumul, tagIndexFactor, bs="fs", m=1), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
# sec
compareML(mod0_3a_3,mod0_3a_2)

# 2. Best structure for year,species,river
mod0 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod1 <- bam(observedLength ~ species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod2 <- bam(observedLength ~ river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

# river expains the most for 1-variable models
AIC(mod0,mod1,mod2) %>% rownames_to_column() %>% arrange(AIC)

# 2-way models
mod3 <- bam(observedLength ~ cohort * species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod4 <- bam(observedLength ~ cohort * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod5 <- bam(observedLength ~ species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

# mod4 is best
AIC(mod0,mod1,mod2,mod3,mod4,mod5) %>% rownames_to_column() %>% arrange(AIC)

# 3-way models
mod6 <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(mod6)

mod6a <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrderedSpecies), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(mod6a)

mod6b <- bam(observedLength ~ cohort * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrderedSpecies), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(mod6b)

#mod6a is best
AIC(mod6,mod6a,mod6b) %>% rownames_to_column() %>% arrange(AIC)

# Any additive 3-way better? NO
mod7 <- bam(observedLength ~ cohort + species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))

mod8 <- bam(observedLength ~ cohort * species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))

mod9 <- bam(observedLength ~ cohort + species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))

AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod7,mod8,mod9) %>% rownames_to_column() %>% arrange(AIC)

summary(mod6)
plot(mod6a)

hist(resid(mod6a))
gam.check(mod6a)

plot(mod6a, shade = TRUE, pages = 2,  seWithMean = TRUE)

#################################
# Predictions
# https://www.fromthebottomoftheheap.net/2017/10/10/difference-splines-i/

mod6a <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

predGrid <- expand.grid(cohortOrderedSpecies = unique(mod6a$model$cohortOrderedSpecies), 
                        river = mod6a$xlevels$river,  
                        ydayCumul = seq(250,1000, by = 30)) %>%
            mutate(cohort = as.numeric(substr(cohortOrderedSpecies,1,4)),
                   species = substr(cohortOrderedSpecies,6,8))

pred <- predict(mod6a, newdata = predGrid)#, type = 'lpmatrix')

predGrid2 <- bind_cols(predGrid,as.data.frame(pred))

ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohort)
```