### 2019-06-28 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
new models

---
title: "Body size patterns"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Plots for the manuscript

```{r Load data, message=FALSE}
library(itsadug)
library(arm)
library(relaimpo)
library(GGally)
library(linkedModels)
library(arrayhelpers)
library(mgcv)
library(tidyverse)
library(gridExtra)

#RECALCULATE gr for dataset. Filtering by maxSampleInterval in the creat ddddG step leaves in trailing lengths from the long samp intervals
# d2 <- d2 %>%
#   group_by(tag) %>%
#   # arrange(tag,sampleNumber) %>%
#   mutate( lagObservedWeight = lead(observedWeight),
#           lagObservedLength = lead(observedLength),
#           grWeight = exp(lagObservedWeight - observedWeight)/as.numeric((lagDetectionDate - detectionDate)),
#           grLength = (lagObservedLength - observedLength)/as.numeric((lagDetectionDate - detectionDate))
#   )  %>%
#   ungroup()

load("D:/projects/linkedModels/data/cd_west_1997.RData")
cd <- cd %>% mutate(observedWeight = ifelse(observedWeight == -9999.9, NA, observedWeight),
                    lagObservedWeight = ifelse(lagObservedWeight == -9999.9, NA, lagObservedWeight))
```

```{r gams for body size - data prep}
speciesIn <- 'bkt'
depVarIn <- 'observedLength'
kIn <- 4 # # of knots for the speces and river smooths

vB <- 
  cd %>% mutate( age = year - cohort,
                 metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1),
                 ydayCumul = yday + 365 * (year - cohort),
                 cf = observedWeight/observedLength^3 * 10000 ) %>%
    filter( ageInSamples %in% 1:9,
            age < 3, 
            year < 2016, area %in% c('inside','trib'), enc == 1
            #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
            )

cohortOffset <- 2006

# 1-way
vB$cohortFactor <- factor(vB$cohort)
vB$cohortOrdered <- factor(vB$cohort, levels = 2000:2013, ordered = T)
#vB$cohortFactorCentered <- vB$cohortFactor - 2006
vB$cohortCentered <- vB$cohort - cohortOffset

vB$speciesFactor <- factor(vB$species, levels = c('bkt','bnt'))
vB$speciesOrdered <- factor(vB$species, levels = c('bkt','bnt'), ordered = T)

vB$riverFactor <- factor(vB$river, levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'))
#vB$riverOrdered <- riverOrdered exists

# 2-way
vB$speciesRiverFactor <- factor(paste(vB$speciesFactor,vB$riverFactor, sep = '_'))
vB$cohortRiverFactor <- factor(paste(vB$cohortFactor,vB$riverFactor, sep = '_'))
vB$seasonRiverFactor <- factor(paste(vB$season,vB$riverFactor, sep = '_'))

vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
vB$speciesCohortOrdered <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
vB$cohortRiverOrdered <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)

# 3-way
vB$cohortSpeciesRiverOrdered <- factor(paste(vB$cohortOrdered,vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)

vB$tagIndexFactor <- factor(vB$tagIndex)

#vB$riverOrdered <- factor(vB$river,
#                         levels=c('west brook', 'wb jimmy', 'wb mitchell', 'wb obear'),
#                         ordered = T)

vB$salmon01 <- ifelse(vB$cohort <= 2004 & vB$river == 'west brook', 1, 0)

vB$speciesGG <- factor(vB$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
vB$seasonGG <- factor(vB$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)

# scaling temperature and flow
vB$meanTemperatureScaled <- as.numeric(scale(vB$meanTemperature))
vB$meanFlowScaled <- as.numeric(scale(vB$meanFlow))
vB$meanTemperatureFlowScaled <- vB$meanTemperatureScaled * vB$meanFlowScaled

seasonalTF <- vB %>% group_by(season) %>%
  summarize(meanTBySeason = mean(meanTemperature, na.rm = TRUE),
            sdTBySeason = sd(meanTemperature, na.rm = TRUE),
            meanFBySeason = mean(meanFlow, na.rm = TRUE),
            sdFBySeason = sd(meanFlow, na.rm = TRUE))

vB <- left_join(vB,seasonalTF) %>%
  mutate(meanTemperatureScaledBySeason = (meanTemperature - meanTBySeason)/sdTBySeason, 
         meanFlowScaledBySeason = (meanFlow - meanFBySeason)/sdFBySeason)

vB$meanTemperatureFlowScaledBySeason <- vB$meanTemperatureScaledBySeason * vB$meanFlowScaledBySeason
```

```{r cohort strength}

cohortStrength <- vB %>%
  distinct(tagIndex, species, metaPop01, cohort) %>% #ignoring 10 fish that were in both o'bear and WB
  group_by(species, metaPop01, cohort) %>%
  summarize(cohortStrength = n()) %>%
  filter(!(cohort < 2000 & species !='ats')) %>% # no trib data
  mutate(cohortStrengthScaled = scale(cohortStrength))

vB <- left_join(vB, cohortStrength)

cohortStrengthWide <- spread(cohortStrength, species, cohortStrengthScaled) %>%
  rename(cohortStrengthScaledBKT = bkt,
         cohortStrengthScaledBNT = bnt,
         cohortStrengthScaledATS = ats
         ) %>%
  dplyr::select(-cohortStrength)

vB <- left_join(vB, cohortStrengthWide)

vB$cohortStrengthScaledBKT <- ifelse(is.na(vB$cohortStrengthScaledBKT), -2.5, vB$cohortStrengthScaledBKT)
vB$cohortStrengthScaledBNT <- ifelse(is.na(vB$cohortStrengthScaledBNT), -2.5, vB$cohortStrengthScaledBNT)
vB$cohortStrengthScaledATS <- ifelse(is.na(vB$cohortStrengthScaledATS), -2.5, vB$cohortStrengthScaledATS)

ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species, color = species)) +
  geom_point(size = 2) +
  geom_line(size = 1.1, color = 'grey40') +
  labs(x = 'Cohort',
       y = 'Cohort strength') +
  theme_publication() +
  facet_grid(~metaPop01)

```

```{r dataIn}
dataIn <- vB %>% filter(cohort %in% 2000:2013, species %in% speciesIn)
```

Run to here to start

```{r raw data graphs}
# number of knots from gam 
kValue <- 9 # from mod6 below, just for graphing

# all fish
ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength)) +
  geom_point(size = 0.5, alpha = 0.1) +
 # geom_smooth(span = 0.1, size = 1.5, se = F) +
  geom_smooth(method='gam', alpha = 0.1, formula=y~s(x, k = kValue), se = F) +
  theme_publication()

# by species
ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength, color = species)) +
  geom_point(size = 0.5, alpha = 0.1) +
 # geom_smooth(span = 0.1, size = 1.5, se = F) +
  geom_smooth(method='gam', alpha = 0.1, formula=y~s(x, k = kValue), se = F) +
 # facet_grid(~species) +
  theme_publication()

# by species, by cohort gridded
# ggplot(vB %>% filter(cohort %in% 2000:2013), aes(ydayCumul,observedLength, group = species, color = (species))) +
#   geom_point(size = 0.5, alpha = 0.1) +
#  # geom_smooth(span = 0.1, size = 1.5, se = F) +
#   geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
#   facet_wrap( ~ cohort) +
#   theme_publication()

# by species (bkt, bnt), by cohort, by river gridded
ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt', 'ats')), aes(ydayCumul,observedLength, group = species, color = (species))) +
  geom_point(size = 0.5, alpha = 0.1) +
 # geom_smooth(span = 0.1, size = 1.5, se = F) +
  geom_smooth(method='gam', alpha=0.1, formula=y~s(x, k = kValue), se = F) +
  facet_grid(riverOrdered ~ cohort) +
  theme_publication()
```

```{r raw data histos}

ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt'), river == "west brook"), aes(observedLength)) +
  geom_freqpoly(aes(color = species)) +
  theme_publication() +
  facet_grid(season ~ year)

# season 3
ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt'), river == "west brook", season == 3), aes(observedLength)) +
  geom_freqpoly(aes(color = speciesGG), binwidth = 6, size = 1.35) +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color='Species') +
  scale_x_continuous("Fork Length (mm)", limits = c(49,200)) +
  scale_y_continuous("Count") +
  #scale_color_manual(values = c("lightblue", "lightgreen")) +
  scale_colour_brewer(palette = "Paired")+
  facet_wrap(~ year)

ggsave('./figures/season3freqpoly.png', width = 11, height = 8, units='in')

# 1 cohort
ggplot(vB %>% filter(cohort == 2009, species %in% c('bkt', 'bnt'), river == "west brook"), aes(observedLength)) +
  geom_freqpoly(aes(color = speciesGG), binwidth = 6, size = 1.35) +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color='Species') +
  scale_x_continuous("Fork Length (mm)", limits = c(49,200)) +
  scale_y_continuous("Count") +
  #scale_color_manual(values = c("lightblue", "lightgreen")) +
  scale_colour_brewer(palette = "Paired")+
  facet_grid(season ~ year)

ggsave('./figures/cohort2009freqpoly.png', width = 11, height = 8, units='in')

```

```{r mean lengths: from MS_Analysis_3spp_WB.Rmd. For Grenada talk}

# Has mean size changed over time?
means <- 
  cd %>% mutate( age = year - cohort ) %>%
    filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
    group_by(species,ageInSamples,year,season,age) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE)
              )

lenSlopes <- means %>%
  group_by( species, ageInSamples, season, age ) %>%
  nest()

lengthModel <- function(df){
  lm( meanLen ~ year, data = df )
}

grModel <- function(df){
  lm( meanGR ~ year, data = df )
}

lenSlopes <- lenSlopes %>%
  mutate( model = map(lenSlopes$data, lengthModel),
          modelGR = map(lenSlopes$data, grModel))

###################################
######### len here, gr below ######
lenSlopesTidy <- lenSlopes %>%
  mutate(glance = map(model, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))

means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +

  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), 
            aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), 
            aes(x=2000,y=195, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +

  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), 
            aes(x=2011.8,y=104, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), 
            aes(x=2011.8,y=74, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+

  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
        legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age, scales='free')

ggsave('./figures/meanSizes.png', width = 11, height = 11, units='in')

# one season/year only
ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 8), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
                legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)
ggsave('./figures/meanSizesAIS_1.png', width = 11, height = 8, units='in')

# is a model with salmon presence absence better than a year model?
means$ats01 <- ifelse(means$year <= 2005, 1, 0)

modMeans1 = lm(meanLen ~ year * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
modMeans2 = lm(meanLen ~ year * factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
modMeans3 = lm(meanLen ~ factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
modMeans4 = lm(meanLen ~ factor(ats01)*year, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
AIC(modMeans1, modMeans2, modMeans3, modMeans4)

# one season/year only
ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(group = factor(ats01)), color = 'black', se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
        legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)

ggsave('./figures/meanSizesAIS_1_ats01.png', width = 11, height = 8, units='in')

# mean increases 
lenSlopesTidy %>%
  filter(species != 'ats') %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
lenSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(lenSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(size = 5) +
  geom_smooth(method='lm', se=FALSE, size = 1.5) +
  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
  labs( x = "Age Season", y = "Slope of mean body length over years" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 40, hjust = 1)) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75)

ggsave(paste0('figures/meanLengthsOverAge.png'), width = 11, height = 10, units='in')

# species diff in increases in length?
sppEffect <- (lm(estimate ~ ageInSamples * (species), data = lenSlopesTidy))
summary(sppEffect)
anova(sppEffect)

# mean diff in slopes btw bkt and bnt
lenSlopesTidy %>%
  filter(species != 'ats') %>%
  group_by(species) %>%
  summarise(means = mean(estimate))

#####################################  
#################################
# growth rate
grSlopesTidy <- lenSlopes %>%
  mutate(glance = map(modelGR, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))

##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanGR, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), 
            aes(x=2000,y=0.68, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), 
            aes(x=2000,y=0.55, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +

  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), 
            aes(x=2011.5,y=0.64, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), 
            aes(x=2011.5,y=0.51, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
        legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG~age)

ggsave('./figures/meanGRs.png', width = 11, height = 11, units='in')

# mean increases 
grSlopesTidy %>%
  filter(species != 'ats') %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
grSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(grSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(size = 5) +
  geom_smooth(method='lm', se=FALSE) +  
  geom_hline(yintercept = 0) +
  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
  labs( x = "Age Season", y = "Slope of mean growth rate over years" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 40, hjust = 1)) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75)

ggsave('./figures/meanGRsOverAge.png', width = 11, height = 11, units='in')

summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))

# species diff in increases in gr?
sppEffectGR <- (lm(estimate ~ ageInSamples * (species), data = grSlopesTidy))
summary(sppEffectGR)
anova(sppEffectGR)

# mean diff in slopes btw bkt and bnt
grSlopesTidy %>%
  filter(species != 'ats') %>%
  group_by(species) %>%
  summarise(means = mean(estimate))

```

```{r size-dep imm or mort? For Spain talk}

### for losses

cd01 <-  cd %>%
  filter(knownZ == 1) %>%
  group_by(tagIndex) %>%
  mutate(first = min(sampleNumber),
         firstObs = sampleNumber == first,
         last = max(sampleNumber),
         lastObs = sampleNumber == last,
         notLastObs = !lastObs,
         nObs = n(),
         age = year - cohort
  ) %>%
  ungroup()

meansByAIS <- cd01 %>%
  group_by(species, ageInSamples, cohort, river) %>%
  summarise(meanLen = mean(observedLength, na.rm =T),
            sdLen = sd(observedLength, na.rm = T),
            n = n())
cd01 <- left_join(cd01, meansByAIS) %>%
  mutate(observedLengthStd = (observedLength - meanLen) / sdLen,
         cohortStd = cohort - 2007)

#cd %>% select(tagIndex,sampleNumber, firstObs,lastObs, knownZ, observedLength) %>% head(50) %>% as.data.frame()
cd01Mod <- cd01 %>% filter(species %in% c('bkt', 'bnt'), river == 'west brook', ageInSamples %in% 1:10, cohort %in% 2000:2014)

ggplot(cd01Mod, aes(observedLengthStd, notLastObs*1, color = factor(species), group = factor(species))) +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) +
 # scale_color_viridis_c() +
  facet_wrap( ~ cohort, scales = 'free')

#### by species and ageInSamples

sensSlopes <- cd01Mod %>%
  group_by( species, ageInSamples ) %>%
  nest()

survModel <- function(df){
  glm(notLastObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
}
enterModel <- function(df){
  glm(firstObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
}

sensSlopes <- sensSlopes %>%
  mutate(modelSurv = map(sensSlopes$data, survModel),
         modelEnter = map(sensSlopes$data, enterModel))

sensSlopesTidy <- sensSlopes %>%
  mutate(surv = map(modelSurv, broom::tidy),
         enter = map(modelEnter, broom::tidy)) %>% 
  unnest(surv, enter, .drop = TRUE) %>%
  filter(term == 'cohortStd') %>% # cohort effect
  rename(estimateSurv = estimate, estimateEnter = estimate1) %>%
  select(species,ageInSamples,estimateSurv, estimateEnter) #%>%
  #gather("type", "estimate", estimateSurv:estimateEnter)

ggplot(sensSlopesTidy, aes(ageInSamples, estimate, color = type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species)

```

```{r dummy graphs for Spain talk}

sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
ggplot(sizesForGraph, aes(year,surv)) +
  geom_line(size = 2) +
  labs(x = 'Year', y = 'Mean body length (mm)') +
  ylim(80,200) +
  theme_publication(base_size = 36)
ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')

survForGraph <- data.frame(size = c(-2,2), surv = c(0.25,0.75))
ggplot(survForGraph, aes(size,surv)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of seen again') +
  theme_publication(base_size = 36)
ggsave('./figures/pSurvivalDummy.png', width = 11, height = 11, units='in')

firstForGraph <- data.frame(size = c(-2,2), surv = c(0.35,0.65))
ggplot(firstForGraph, aes(size,surv)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of first observation') +
  theme_publication(base_size = 36)
ggsave('./figures/pFirstDummy.png', width = 11, height = 11, units='in')

sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
ggplot(sizesForGraph, aes(year,surv)) +
  geom_line(size = 3) +
  labs(x = 'Year', y = 'Mean body length (mm)') +
  ylim(80,200) +
  theme_publication(base_size = 36)
ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')

#### show interactions

survForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.52,0.55,0.51,0.24,0.54,0.82))
ggplot(survForGraph, aes(size,surv, group = ind)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of seen again') +
  theme_publication(base_size = 36)
ggsave('./figures/pSurvivalDummy3.png', width = 11, height = 11, units='in')

firstForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.42,0.45,0.41,0.14,0.44,0.72))
ggplot(firstForGraph, aes(size,surv, group = ind)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of first observation') +
  theme_publication(base_size = 36)
ggsave('./figures/pFirstDummy3.png', width = 11, height = 11, units='in')

```

```{r get betas from growth model for sensitivities}
load("D:/projects/linkedModels/data/out/allBetas.RData") #loads df 'q'
qFig1 <- q %>% filter(riverGG == 'WB', species %in% c('bkt','bnt'))
qFig1$season <- as.numeric(as.character(qFig1$season))
qFig1$isYOY <- as.numeric(as.character(qFig1$isYOY))
qFig <- qFig1 %>%
  mutate( varText = tolower(betaGG)) %>%
  mutate( varText = ifelse(varText == 'temp*flow','temp:flow', varText)) %>%
  mutate( varText = ifelse(varText == 'bnt*bkt','bkt:bnt', varText)) %>%
  select(isYOY,season,species,median,varText)
qFig$seasonGG <- factor(qFig$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
#get slopes of env variables

load('D:/projects/linkedModels/data/out/flowSlopesTidy.RData')
flowSlopesTidy2 <- flowSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'flow')
load('D:/projects/linkedModels/data/out/tempSlopesTidy.RData')
tempSlopesTidy2 <- tempSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'temp')
load('D:/projects/linkedModels/data/out/nSlopesTidy.RData')
nSlopesTidy2 <- nSlopesTidy %>% select(estimate, season, species) %>% mutate(varText = species)
nSlopesTidy2$seasonGG <- factor(nSlopesTidy2$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
nSlopesTidy2 <- nSlopesTidy2 %>% select(estimate,seasonGG,varText)

allSlopesTidy <- bind_rows(nSlopesTidy2, tempSlopesTidy2, flowSlopesTidy2)

allSlopesTidyWBetas <- left_join(qFig, allSlopesTidy, by = c('seasonGG', 'varText')) %>% mutate(sens = median * estimate)

ggplot(allSlopesTidyWBetas%>%filter(!is.na(sens), varText != 'ats'), aes(seasonGG, sens, color = varText)) +
  geom_point() +
  facet_grid(species + isYOY ~ varText, scales = 'free')
```

```{r }
############# 1st attempt at imm em
mod1 <- glm(notLastObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
#mod2 <- glm(notLastObs*1 ~ observedLengthStd * factor(cohortStd) * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
mod3 <- glm(notLastObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)

AIC(mod1,mod3)

summary(mod1)
anova(mod1, test="Chi")

pred <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))

pred$pred <- predict.glm(mod1, newdata = pred, type = 'response')

u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
pred <- left_join(pred, u)

pred$seasonGG <- factor(pred$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
pred$cohort <- pred$cohortStd + 2007
pred$speciesGG <- factor(pred$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(pred, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_continuous("Probability of survival") +
  facet_grid(species ~ ageInSamples, scales = 'free')

ggplot(pred %>% filter(cohortStd %in% 0, age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
  geom_point(size = 2) +
  scale_x_continuous("Standardized length") +
  scale_y_continuous(lim = c(0,1),"Probability of seen again") +
  theme_publication(base_size = 28) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal"
        ) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)

ggsave('./figures/pSurvival.png', width = 11, height = 11, units='in')

################################
# gains - immigration, firstObs

ggplot(cd01Mod, aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) +
 # scale_color_viridis_c() +
  xlim(c(-2.75,2.75)) +
  facet_grid(ageInSamples ~ cohort)

ggplot(cd01Mod%>%filter(ageInSamples == 3), aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) +
 # scale_color_viridis_c() +
  xlim(c(-2.75,2.75)) +
  facet_wrap( ~ cohort)

modi1 <- glm(firstObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
modi2 <- glm(firstObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
modi3 <- glm(firstObs*1 ~ observedLengthStd * species , family = 'binomial', data = cd01Mod)
modi4 <- glm(firstObs*1 ~ observedLengthStd * factor(ageInSamples), family = 'binomial', data = cd01Mod)
modi5 <- glm(firstObs*1 ~ observedLengthStd, family = 'binomial', data = cd01Mod)

AIC(modi1,modi2,modi3,modi4,modi5) %>% rownames_to_column() %>% arrange(AIC)

summary(modi1)
anova(modi1, test="Chi")
###
predi <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))

predi$pred <- predict.glm(modi1, newdata = predi, type = 'response')

u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
predi <- left_join(predi, u)

ggplot(predi, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_continuous("Probability of first observation") +
  facet_grid(species ~ ageInSamples, scales = 'free')

predi$seasonGG <- factor(predi$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
predi$cohort <- predi$cohortStd + 2007
predi$speciesGG <- factor(predi$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(predi %>% filter(cohortStd %in% c(0), species == 'bkt', age < 3), aes(observedLengthStd, pred)) +
  geom_point() +
  scale_color_viridis_c() +
  scale_x_continuous("Standardized length") +
  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
  theme_publication() +
  facet_grid(seasonGG ~ age)

ggplot(predi %>% filter(cohortStd %in% c(0), age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
  geom_point(size = 2) +
  scale_x_continuous("Standardized length") +
  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
  theme_publication(base_size = 28) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal"
        ) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)

ggsave('./figures/pFirstObs.png', width = 11, height = 11, units='in')
```

```{r gam models - find best lm, before fitting detailed gams - only need to run for model selection}
# only including river, season, salmon01 becaseu we assume these have categorical effects

mod0 <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod1 <- bam(observedLength ~ riverFactor + factor(season) + factor(salmon01) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod2 <- bam(observedLength ~ riverFactor * factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod3 <- bam(observedLength ~ riverFactor + factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod4 <- bam(observedLength ~ riverFactor + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod5 <- bam(observedLength ~ factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

AIC(mod0,mod1,mod2,mod3,mod4,mod5,mod6) %>% rownames_to_column() %>% arrange(AIC)
summary(mod0)
plot(mod0)
gam.check(mod0)

```

```{r gam structure given best model from above}

# BKT, mod0 is best fixed effects model

system.time(
mod0_YDCcr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCc_r <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) , 
               data = dataIn, 
               method = "REML")
)

AIC(mod0_YDCcr, mod0_YDCc_r) %>% rownames_to_column() %>% arrange(AIC)

system.time(
mod0_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_Cr_CSr_Ts_Fs_TFs <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = season) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = season) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = season), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)

system.time(
mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_C_CS_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)
summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)

# take out s()'s

system.time(
mod0_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_Cr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               #s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
              # s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
             #  s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
             #  s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)
########

AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CSr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_Tsr_Fsr_TFsr,mod0_YDCcr_CSr_Tsr_Fsr_TFsr, mod0_YDCcr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)

#                           rowname        df      AIC
# 1  mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr 277.67003 401304.4
# 2   mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr 281.22421 401407.3
# 3   mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr 277.09498 401772.0
# 4     mod0_YDCcr_CSr_Tsr_Fsr_TFsr 279.84632 402037.9
# 5    mod0_YDCcr_C_CS_Tsr_Fsr_TFsr 269.69828 402073.1
# 6     mod0_YDCcr_Cr_CSr_Tr_Fr_TFr 207.97536 402187.1
# 7     mod0_YDCcr_Cr_CSr_Ts_Fs_TFs 190.64645 402464.7
# 8      mod0_YDCcr_Cr_Tsr_Fsr_TFsr 289.33427 403529.7
# 9         mod0_YDCcr_Tsr_Fsr_TFsr 283.93242 404139.1
# 10              mod0_YDCcr_Cr_CSr 184.24038 413605.9
# 11                     mod0_YDCcr 168.77693 420503.0
# 12                    mod0_YDCc_r  71.99108 422677.4

summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)

```

```{r predictions}
# BKT
# system.time(
# mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
#                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
#                s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
#                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
#                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
#                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
#                data = dataIn, 
#                method = "REML")
# )

predGridBKT <- expand.grid(cohortRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$cohortRiverFactor),
                           seasonRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$seasonRiverFactor),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>% #seq(-1.5,1.5, by = 0.5)) %>%
                          # meanTemperatureFlowScaledBySeason = seq(-1.5,1.5, by = 0.5)) %>%
                    mutate(cohortFactor = as.numeric(substr(cohortRiverFactor,1,4)),
                           riverFactor = substring(cohortRiverFactor,6),
                           season = substring(seasonRiverFactor,1),
                           #cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           #seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           #cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

ydayCumulSeason <- distinct(dataIn, ydayCumul,season,age) %>% arrange(ydayCumul) %>% mutate(age_season = paste(age,season,sep = '_')) %>%
  group_by(age_season) %>% summarise(age=min(age),season=min(season),minY=min(ydayCumul),maxY=max(ydayCumul))
#38 obear fish were captured in january, but get season 4 and are age 1, so have very small ydayCumul -need to manually reset 
#dataIn %>% filter(age==1,season==4,ydayCumul==367) %>% select(tag,cohort,detectionDate,year,ydayCumul,yday,river,age,season)
ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '1_4',670,ydayCumulSeason$minY)
ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '2_4',1035,ydayCumulSeason$minY)
ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '0_3',0,ydayCumulSeason$minY)
ydayCumulSeason %>% mutate(minY2 = lead(minY))

yDayAll <- data.frame(y = 250:1000)
# pick up here
#### need to do something like this  left_join(yDayAll, ydayCumulSeason, by = c('yDay' = 'minY'))

predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                           cohortFactor = unique(dataIn$cohort),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., ydayCumulSeason) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

predBKT <- predict(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, newdata = predGridBKT)#, type = 'lpmatrix')

predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')

ggplot(predGrid2BKT %>% filter(salmon01 == 0),

       aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(~riverFactor, scales = 'free_y')

```

```{r}
#earlier runs before setting fixed effect first. kept forf reference
# BKT 
#             rowname        df      AIC
# 1     modCR_SR_c_cS 162.34756 414267.2
# 2      modCR_S_c_cS 163.43154 414602.7
# 3        modCR_c_cS 161.63691 414750.3
# 4  modCR_SR_c_cSR3a  88.21646 415388.9
# 5  modCR_SR_c_cSR3b 167.32968 416062.3
# 6    modCR_SR_c_cSR 163.93920 416587.5
# 7   modCR_SR_c_cSR3 173.36884 416611.4
# 8         modCR_S_c 157.23995 418645.9
# 9           modCR_c 158.50186 418665.1
# 10            modCR 154.36186 419135.1
# 11           modC_R  63.10879 421258.1
# 12              mod  17.99204 427309.6
# 13             modC  48.73277 431440.2
# 14             modR  22.80483 432214.7

# BNT
#             rowname       df      AIC
# 1  modCR_SR_c_cSR3a 89.52290 231873.4
# 2  modCR_SR_c_cSR3b 89.54674 231873.6
# 3     modCR_SR_c_cS 84.40505 231935.7
# 4      modCR_S_c_cS 83.47382 231962.5
# 5   modCR_SR_c_cSR3 88.15840 232100.8
# 6        modCR_c_cS 80.92171 232142.0
# 7    modCR_SR_c_cSR 80.89406 232201.5
# 8         modCR_S_c 79.53048 232281.7
# 9           modCR_c 77.80808 232420.2
# 10            modCR 78.58400 233816.0
# 11           modC_R 53.21199 234214.3
# 12              mod 13.99759 238544.8
# 13             modR 14.97078 243472.0
# 14             modC 48.73277 431440.2

```

```{r best models for each species}

# Brook trout

#using salmon01 * river because we need diff effects per river - even though AIC is worse for this model
modBKT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn),
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bkt'), method = "REML")

summary(modBKT)

#plot(modBKT, pages = 8, scale = 0)

predGridBKT <- expand.grid(cohortRiverOrdered = unique(modBKT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaledBKT = seq(-2,2, by = 0.1)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predBKT <- predict(modBKT, newdata = predGridBKT)#, type = 'lpmatrix')

predGrid2BKT <- bind_cols(predGridBKT, as.data.frame(predBKT)) %>% mutate(pred = predBKT, species = 'bkt')

#################################
# Brown trout

modBNT <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaledBKT, bs = 'cr', k = kIn, by = metaPop01) +
               s(cohortStrengthScaledBNT, bs = 'cr', k = kIn) +
               s(cohortStrengthScaledATS, bs = 'cr', k = kIn), 
             data = vB %>% filter(cohort %in% 2000:2013, species %in% 'bnt'), method = "REML")

summary(modBNT)
#plot(modBNT, pages = 8, scale = 0)

predGridBNT <- expand.grid(cohortRiverOrdered = unique(modBNT$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           metaPop01 = c(0,1),
                           cohortStrengthScaledBKT = seq(-2,2, by = 0.25),
                           cohortStrengthScaledBNT = seq(-2,2, by = 0.25),
                           cohortStrengthScaledATS = seq(-2,2, by = 0.25)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predBNT <- predict(modBNT, newdata = predGridBNT)#, type = 'lpmatrix')

predGrid2BNT <- bind_cols(predGridBNT, as.data.frame(predBNT)) %>% mutate(pred = predBNT, species = 'bnt')

##################
# combine species' predictions

predGrid2Both <- bind_rows(predGrid2BKT, predGrid2BNT) %>% dplyr:::select(-predBKT, -predBNT)

predGrid2Both$riverGG <- factor(predGrid2Both$riverOrdered,
                         levels=c('west brook', 'wb jimmy', 'wb mitchell',"wb obear"),
                         labels = c("West Brook","Open Large","Open Small","Isolated Small"), ordered = T)
predGrid2Both$speciesGG <- factor(predGrid2Both$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

save(modBKT,modBNT,predBKT,predBNT,predGrid2Both, file = paste0('./data/out/bestModsBKTBNT.RData'))

```

```{r graphs of best models}
# load('./data/out/bestModsBKTBNT.RData')
source('./functions/plotGams.R')

# cohorts by river/species
ggplot(predGrid2Both %>% 
         filter(
         (cohortStrengthScaledBKT == 0 & salmon01 == 0) | (cohortStrengthScaledBKT == 0 & cohortStrengthScaledBNT == 0 & cohortStrengthScaledATS == 0 & metaPop01 == 1 & salmon01 ==0)),
#           cohortStrengthScaledBKT == 0, cohortStrengthScaledBNT == 0, cohortStrengthScaledATS == 0, metaPop01 == 1, salmon01 == 0),
       aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(speciesGG ~ riverGG, scales = 'free_y')

ggsave('./figures/predLengths.png', width = 11, height = 11, units='in')

# specific splines by species
(plotCohort <- plotGamVarBySpecies(var = 's(cohort)', xAxisLab = 'Cohort'))
(plotCohortStrength <- plotGamVarBySpecies(var = 's(cohortStrengthScaled)', xAxisLab = 'Cohort Strength'))

#
plot.gam(modBKT, pages = 10)

# main spline over ydayCumul
plotBKT = plot.gam(modBKT, select = 1, scale = 0)
plotBNT = plot.gam(modBNT, select = 1, scale = 0)

plotBKTydayCumul <- data.frame(x = plotBKT[[1]]$x, y = plotBKT[[1]]$fit, se = plotBKT[[1]]$se) %>% mutate(species = 'bkt')
plotBNTydayCumul <- data.frame(x = plotBNT[[1]]$x, y = plotBNT[[1]]$fit, se = plotBNT[[1]]$se) %>% mutate(species = 'bnt')
plotBothydayCumul <- bind_rows(plotBKTydayCumul,plotBNTydayCumul) %>% mutate(upperCI = y + se * 1.96, lowerCI = y - se * 1.96)

plotBothydayCumul$speciesGG <- factor(plotBothydayCumul$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(plotBothydayCumul, aes(x, group = speciesGG)) + 
  geom_ribbon(aes( ymin = lowerCI, ymax = upperCI, fill  = speciesGG), alpha = 0.5) +
  geom_line(aes(y = y, linetype = speciesGG), size = 1.5) +
  scale_fill_manual(values = c('grey30', 'grey70')) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       linetype = 'Species') +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face="plain"))

# cohort effects

ggplot(tmp2, aes(x,y)) + geom_line()

appraise(modBKT)

```

```{r early gams}
# some early exploration

# 1. Do we need separate splines on cohort in s()? YES, cohortOrdered is best
# 
mod0_0 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_1 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr', by = (cohort)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_2 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr', by = (cohortFactor)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_3 <- bam(observedLength ~ 
                cohort + 
                s(ydayCumul, bs = 'cr') + 
                s(ydayCumul, bs = 'cr', by = (cohortOrdered)), 
              data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_4 <- bam(observedLength ~ cohort + 
                s(ydayCumul, bs = 'cr') + 
                s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)),
                data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

#mod0_4 is best
AIC(mod0_0,mod0_1,mod0_2,mod0_3,mod0_4) %>% rownames_to_column() %>% arrange(AIC)

# 2. structure of smooths
# with ordered factor, need to include an overall smooth, http://www.sfs.uni-tuebingen.de/~jvanrij/Tutorial/GAMM.html 
mod0_4a <- gam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod0_4b <- gam(observedLength ~ cohortOrdered + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

# mod0_3b is best, but we want to estimate a trend as in mod0_3a
# add + s(ydayCumul, bs = 'cr') to models
AIC(mod0_4,mod0_4a,mod0_4b) %>% rownames_to_column() %>% arrange(AIC)

# 3. gam vs. bam for this big dataset? bam() is MUCH faster!
system.time( mod0_4a_1 <- gam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
# 316 sec
summary(mod0_3a_1)

system.time( mod0_4a_2 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
# 8 sec
summary(mod0_4a_2)

# 4. bs = 'cr', vs default 'tp'. AIC for 'cr' is ~ 300 lower. Stick with 'cr'
mod0_4a_2a <- bam(observedLength ~ cohort + s(ydayCumul) + s(ydayCumul, by = (cohortOrderedSpecies)), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

AIC(mod0_4a_2,mod0_4a_2a)

# 5. Include Individual as a RE??
# Taking a long time, maybe run overnight
vB$tagIndexFactor <- factor(vB$tagIndex)
system.time( mod0_3a_3 <- bam(observedLength ~ year + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = (yearOrdered)) + s(ydayCumul, tagIndexFactor, bs="fs", m=1), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML") )
# sec
compareML(mod0_3a_3,mod0_3a_2)

# 2. Best structure for year,species,river
mod0 <- bam(observedLength ~ cohort + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod1 <- bam(observedLength ~ species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod2 <- bam(observedLength ~ river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

# river expains the most for 1-variable models
AIC(mod0,mod1,mod2) %>% rownames_to_column() %>% arrange(AIC)

# 2-way models
mod3 <- bam(observedLength ~ cohort * species + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod4 <- bam(observedLength ~ cohort * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

mod5 <- bam(observedLength ~ species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

# mod4 is best
AIC(mod0,mod1,mod2,mod3,mod4,mod5) %>% rownames_to_column() %>% arrange(AIC)

# 3-way models
mod6 <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(mod6)

mod6a <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrderedSpecies), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(mod6a)

mod6b <- bam(observedLength ~ cohort * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrderedSpecies), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")
summary(mod6b)

#mod6a is best
AIC(mod6,mod6a,mod6b) %>% rownames_to_column() %>% arrange(AIC)

# Any additive 3-way better? NO
mod7 <- bam(observedLength ~ cohort + species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))

mod8 <- bam(observedLength ~ cohort * species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))

mod9 <- bam(observedLength ~ cohort + species + river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr', by = cohortOrdered), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')))

AIC(mod1,mod2,mod3,mod4,mod5,mod6,mod7,mod8,mod9) %>% rownames_to_column() %>% arrange(AIC)

summary(mod6)
plot(mod6a)

hist(resid(mod6a))
gam.check(mod6a)

plot(mod6a, shade = TRUE, pages = 2,  seWithMean = TRUE)

#################################
# Predictions
# https://www.fromthebottomoftheheap.net/2017/10/10/difference-splines-i/

mod6a <- bam(observedLength ~ cohort * species * river + s(ydayCumul, bs = 'cr') + s(ydayCumul, bs = 'cr'), data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt','bnt')), method = "REML")

predGrid <- expand.grid(cohortOrderedSpecies = unique(mod6a$model$cohortOrderedSpecies), 
                        river = mod6a$xlevels$river,  
                        ydayCumul = seq(250,1000, by = 30)) %>%
            mutate(cohort = as.numeric(substr(cohortOrderedSpecies,1,4)),
                   species = substr(cohortOrderedSpecies,6,8))

pred <- predict(mod6a, newdata = predGrid)#, type = 'lpmatrix')

predGrid2 <- bind_cols(predGrid,as.data.frame(pred))

ggplot(predGrid2, aes(ydayCumul,pred, color = species)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohort)
```